A Comparison of Parallel Graph Processing Benchmarks

### 2017-02-17 papers/publication/paper-pollard.tex
Updated some figures and caption text.

Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} belies the actual variance in results. Each platform in Fig.~\ref{fig:pr} has a standard deviation comparable to that of

%> sd(pr_time$Time[pr_time$Sys == "PowerGraph"])
%[1] 0.4654061
%> sd(sssp_time$Time[sssp_time$Sys == "PowerGraph"])
%[1] 0.1475608
%> sd(pr_time$Time[pr_time$Sys == "GAP"])
%[1] 0.01957317
%> sd(sssp_time$Time[sssp_time$Sys == "GAP"])
%[1] 0.01608644
%> sd(pr_time$Time[pr_time$Sys == "GraphMat"])
%[1] 0.9305763
%> sd(sssp_time$Time[sssp_time$Sys == "GraphMat"])
%[1] 0.02191714
%> sd(pr_time$Time[pr_time$Sys == "GraphBIG"])
%[1] 3.308351
%> sd(sssp_time$Time[sssp_time$Sys == "GraphBIG"])
%[1] 0.2165655

\begin{figure}
	\centering
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/pr_time.pdf}
	\end{minipage}
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/pr_iters.pdf}
	\end{minipage}
	\caption{The $y$-axis is logarithmic only for the left figure. GraphMat continues to run until none of the vertices' ranks change. For the others, we use the stopping condition that the sum of the changes in the weights is no more than $6 \times 10^{-8}$, or approximately machine epsilon for single precision floating point numbers.}
	\label{fig:pr}
\end{figure}
\todo{Check this again---that's a huge discrepancy!}

@@

In Fig.~\ref{fig:bfs-scaling}, we see BFS scalability\dots

\begin{figure}
	\centering
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1.}
	\label{fig:bfs-scaling}
\end{figure}
\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}

@@

RAPL also allows the measurement of DRAM power and energy, whose results are shown in Fig.~\ref{fig:ram-power}.

\begin{figure}
	\centering
	\includegraphics[width=0.6\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
	\caption{RAM Power Consumption, averaged across all roots. [TODO: Explain more]}
	\label{fig:ram-power}
\end{figure}

\section{Future Work and Conclusion}
The problem of choosing a system for graph processing at large scale is far from being simple. We make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation. Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.

>>>>

Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} compresses the apparent variance in runtime. Each platform in Fig.~\ref{fig:pr} has a relative standard deviation between $1/4$ and $1/2$ that of the same system executing SSSP.

\begin{figure}
	\centering
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth, trim=0pt 18pt 18pt 0pt, clip]{graphics/pr_time.pdf}
	\end{minipage}
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 0pt, clip]{graphics/pr_iters.pdf}
	\end{minipage}
	\caption{The $y$-axis is logarithmic only for the left figure. GraphMat continues to run until none of the vertices' ranks change. For the others, we use the stopping condition that the sum of the changes in the weights is no more than $6 \times 10^{-8}$, or approximately machine epsilon for single precision floating point numbers.}
	\label{fig:pr}
\end{figure}
\todo{Check this again---that's a huge discrepancy!}

@@

Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time and $n$ is the number of threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, though GraphBIG has the best scaling. One explanation is GraphBIG is the slowest system so there is more room for improvement. Additionally, a scale of 20 generates a small graph by today's standards and thus the focus is on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
\begin{figure}
	\centering
	\begin{minipage}{0.9\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic. The exception is the extra space at 72 threads for readability. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}
\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}

@@

RAPL also allows the measurement of DRAM power and energy, whose results are shown in Fig.~\ref{fig:power}.

\todo{Make text bigger}
\begin{figure}
	\centering
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_ram_power.pdf}
	\end{minipage}
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth, trim=0 36pt 18pt 0, clip]{graphics/bfs_cpu_power.pdf}
	\end{minipage}
	\caption{Similar to Fig.~\ref{fig:bfs-time} we plot RAM and CPU Power Consumption for each of the 32 roots. Since the Graph500 runs multiple roots per execution, we only get a single datapoint. The baseline was computed by monitoring power consumption while a program containing only one call to the C \texttt{unistd} library function \texttt{sleep(10)} (ten seconds).}
	\label{fig:power}
\end{figure}

\section{Future Work and Conclusion}
We have presented an approach to analyze parallel graph processing performance at fine detail, both measuring power consumption and the two fundamental phases of execution. Still, the problem of choosing a system for graph processing at large scale is far from being simple. We make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation. Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.
